Beyond Lip Reading - Inferring Voice Qualities from Visual Data

### 2024-04-08 PaperForReview.tex
Additional work on 2024-04-08

There have been no shortages of papers and processes for distilling spoken words from visual-only data sources, i.e. performing lip reading. Fenghour \etal 2021 \cite{Fenghour2021} does a thorough job reviewing the corpuses, pre-processing techniques, neural network architectures, and results of these processes. Ultimately, the target goal is to produce understandable words / sentences for consumption. However, there are many other complexities to human vocal communication which these methods are not trained to reproduce. As a simple example, rising and falling intensity of speaking voice can be suggestive of the certainty of the speaker to certain facts, or rising pitch can suggest if the speaker is posing the statement as a question or topic of discussion versus a fact.

@@

Our methodology for performing this research was extensive due to the nature of the problem - specifically, we needed a corpus that would represent both a speaker's baseline audio characteristics (pitch / intensity / harmonics / etc.) as well as their audio characteristics when they deviate from that baseline. After all, as we speak, the qualities of our voice modulate due to changes in emotional state and as we purposefully modulate our voice to convey meaning and context such as excitement, boredom, sarcasm, amongst other information. To achieve this, we developed our own corpus based on long-form interviews conducted in a semi-consistent environment. The long-form nature of the source datasets would facilitate 1) having enough data for an individual speaker to infer a baseline and 2) provide ample time for the conversation to inspire a diversity of emotions that would cause the speaker to deviate from their baseline speaking qualities. We then collected the necessary metadata from the models and

@@

@@

>>>>

There have been no shortages of papers and processes for distilling spoken words from visual-only data sources, i.e. performing lip reading. Fenghour \etal 2021 \cite{Fenghour2021} does a thorough job reviewing the corpora, pre-processing techniques, neural network architectures, and results of these processes. Ultimately, the target goal is to produce understandable words / sentences for consumption. However, there are many other complexities to human vocal communication which these methods are not trained to reproduce. As a simple example, a rising pitch can suggest if the speaker is posing the statement as a question or topic of discussion versus a fact. After all, as we speak, the qualities of our voice modulate due to changes in emotional state and as we purposefully modulate our voice to convey meaning and context such as excitement, boredom, sarcasm, amongst other information.

@@

Our methodology for performing this research was extensive due to the nature of the problem - specifically, we needed a corpus that would represent both a speaker's baseline audio characteristics (pitch / intensity / harmonics / etc.) as well as their audio characteristics when they deviate from that baseline.

To achieve this, we developed our own corpus based on long-form conversations conducted in a semi-consistent environment. The long-form nature of the source datasets would facilitate 1) having enough data for an individual speaker to infer a baseline and 2) provide ample time (and stimulation) for the conversation to inspire a diversity of emotions that would cause the speaker to deviate from their baseline speaking qualities. We then collected the necessary metadata from the models using standard speech analysis tooling and libraries to create the speaker diarization and ultimately the audio metadata. We then

@@

As our "dynamic" data set, we chose the "Joe Rogan Experience Podcast", available as MP4 audio / video files on archive.org. We chose this due to the numeracy of the episodes, the diversity of speakers, the length of each episode (with many eclipsing 2 hours), and the dynamism of the host, Joe Rogan. We pulled select episodes from 200 to 600, limited to episodes with 1 guest which did not span multiple episodes, and also removed conversations where the speaker was not visible speaking, as was the case the the John McAfee episode.

@@

Once we had the mp4 files, we extracted the video and audio into separate files. For the audio files, we performed speaker diarization (i.e. split the audio file into segments by the speaker) using the pyannote.audio library\cite{Bredin23}. We then fed these segmented audio files into a Pratt\cite{Pratt}-based library called parselmouth\cite{Parselmouth}.  Parselmouth is a convenient python-based tool we used to extract the voice qualities we were looking for, in 0.1 second intervals, which would become the basis for the data points our model would attempt to predict. We also accumulated this data to create speaker median values, which we would predict in tandem with the temporally-associated values. Out hope was that even if the model fails to correctly predict the median values, we could see predictions that were correctly above or below the median when relevant.
